---
title: "P8105 HOMEWORK 1"
author: "Anu Singh"
date: "2025-09-20"
output: github_document
---

## PROBLEM 1

Loading the required library, and displaying portion of dataset below.
```{r}
library("moderndive")
data("early_january_weather")
head(early_january_weather)
```

The early_january_weather dataset contains hourly meteorological data from Newark Liberty International Airport (EWR) for January 1st to 15th in 2013. While the R help documentation mentions that the dataset includes data from three NYC airports (EWR, LGA, and JFK), a quick look at the origin column using unique() reveals only EWR data is present.

The dataset includes `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns, with the key variables/measurements including:

* temp - temperature in degrees Fahrenheit
* dewp - dewpoint in degrees Fahrenheit
* humid - relative humidity (percentage)
* wind_dir - wind direction in degrees
* wind_speed - wind speed in miles per hour
* wind_gust - wind gust speed in miles per hour
* precip - precipitation in inches
* pressure - sea level pressure in millibars
* visib - visibility in miles
* time_hour - date and hour of recording (POSIXct format)

The mean temperature during January 2013 was `r round(mean(early_january_weather$temp), 2)` degrees Fahrenheit.

Using ggplot() from the tidyverse library, a scatterplot of time vs. temperature was created.

```{r scatterplot, message = FALSE}
library(tidyverse)
scatterplot_temperature = ggplot(early_january_weather, 
       aes(x = time_hour, y = temp, color = humid )) + 
	geom_point() +
  labs(title = "Temperature vs. Time (with Humidity)",
       x = "Date and Hour",
       y = "Temperature (°F)",
       color = "Relative Humidity (%)" )
scatterplot_temperature
```

From observing the scatterplot, an upward trend is apparent, as days progressed from early January to mid-January, the temperature generally increased.An oscillating pattern is also visible from the plot, where temperatures typically rise during daytime, and fall during night. There is no distinct trend with humidity, but it appears that humidity decreases at higher temperatures (noticeable at peaks in the graph).

The plot created has been saved in the project directory as a PNG file.
```{r graph_PNG, message = FALSE}
ggsave("scatterplot_temperature.png", plot = scatterplot_temperature)
```

## PROBLEM 2

```{r}

```

