---
title: "P8105 HOMEWORK 1"
author: "Anu Singh"
date: "2025-09-20"
output: github_document
---

## PROBLEM 1

I loaded the required library, and a portion of the dataset is displayed below.
```{r}
library("moderndive")
data("early_january_weather")
head(early_january_weather)
```

The early_january_weather dataset contains hourly meteorological data from Newark Liberty International Airport (EWR) for January 1st to 15th in 2013. While the R help documentation mentions that the dataset includes data from three NYC airports (EWR, LGA, and JFK), a quick look at the origin column using unique() reveals only EWR data is present.

The dataset includes `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns, with the key variables/measurements including:

* temp - temperature in degrees Fahrenheit
* dewp - dewpoint in degrees Fahrenheit
* humid - relative humidity (percentage)
* wind_dir - wind direction in degrees
* wind_speed - wind speed in miles per hour
* wind_gust - wind gust speed in miles per hour
* precip - precipitation in inches
* pressure - sea level pressure in millibars
* visib - visibility in miles
* time_hour - date and hour of recording (POSIXct format)

The mean temperature during January 2013 was `r round(mean(early_january_weather$temp), 2)` degrees Fahrenheit.

Using ggplot() from the tidyverse library, a scatterplot of time vs. temperature was created.

```{r scatterplot, message = FALSE}
library(tidyverse)
scatterplot_temperature = ggplot(early_january_weather, 
       aes(x = time_hour, y = temp, color = humid )) + 
	geom_point() +
  labs(title = "Temperature vs. Time (with Humidity)",
       x = "Date and Hour",
       y = "Temperature (°F)",
       color = "Relative Humidity (%)" )
scatterplot_temperature
```

From observing the scatterplot, an upward trend is apparent, as days progressed from early January to mid-January, the temperature generally increased.An oscillating pattern is also visible from the plot, where temperatures typically rise during daytime, and fall during night. There is no distinct trend with humidity, but it appears that humidity decreases at higher temperatures (noticeable at the small peaks in the graph).

The plot created has been saved in the project directory as a PNG file.
```{r graph_PNG, message = FALSE}
ggsave("scatterplot_temperature.png", plot = scatterplot_temperature)
```

## PROBLEM 2

Using data.frame() a dataframe has been generated below, as per the given requirements.
```{r}
# setting seed to produce same random values each run
set.seed(7923)

problem2_df = data.frame(
  # random sample of 10 from standard Normal dist.
  numeric_var = rnorm(10), 
  
  # logical vector (TRUE if > 0)                             
  logical_var = rnorm(10) > 0, 
  
  # character vector of length 10
  character_var = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),  
  
  # factor vector of length 10,  with 3 levels
  factor_var = factor(rep(c("low", "medium", "high"), length.out = 10))  
)

problem2_df
```
After generating the above dataframe, the mean for each variable will be derived below.

```{r}
mean(problem2_df %>% pull(numeric_var))
mean(problem2_df %>% pull(logical_var))
mean(problem2_df %>% pull(character_var))
mean(problem2_df %>% pull(factor_var))
```
After running the above lines of code, only means for the numeric and logical variables were calculated. The mean calculations for the character and factor variables resulted in an error.

I will now apply as.numeric() to logical, character, and factor variables, and see what happens.
```{r as_numeric, eval = FALSE}
as.numeric(problem2_df %>% pull(logical_var))
as.numeric(problem2_df %>% pull(character_var))
as.numeric(problem2_df %>% pull(factor_var))
```

The following is observed:

* Logical variable - TRUE values are converted to 1 and FALSE are converted to 0
* Character variable - NAs and a warning is produced, this is because letters cannot be converted to numbers
* Factor variable - Converts the factors to their underlying numeric values (2 = low, 3 = medium, 1 = high). The assigned values represent the factor levels

The conversion of the above value explains the results for means we saw earlier. The mean of the logical variable generates the proportion of TRUE values (hence we saw 0.6). Whereas, the means of character and factor variables result in an error because there is no way to convert letters into meaningful numeric values, and averaging categorical variables such as low, medium, high is meaningless and does not make sense. 


